\documentclass{ximera}
\title{HW 1}

\begin{document}

    \begin{abstract}  Due April 10.  \end{abstract}
    \maketitle



\begin{exercise}
    Let $M$ be an $n$-dimensional topological manifold and charts $(U, \varphi )$ and $(V, \psi )$. Assume there is a collection of charts $\{ (U_i, \varphi _i ) \} _{i \in I} $  such that for each $i \in I$, the chart $(U_i, \varphi_i ) $ is compatible with both $(U, \varphi) $ and $(V, \psi)$. If 
    \[    U \cap V \subset \bigcup_{i \in I} U_i ,  \]
    then $U$ and $V$ are compatible. 

    \begin{remark}
        Before, I claimed that compatibility of charts was an equivalence relation. This is false. For example, the ``empty chart'' $(\emptyset , \emptyset )$ is compatible with all charts!
    \end{remark}
\end{exercise}


\begin{exercise}
    Let $M$ be an $n$-dimensional topological manifold and $\mathcal{A}$ a smooth atlas. Show that there is a unique smooth structure $\mathcal{S}$ containing $\mathcal{A}$. Moreover, show that $\mathcal{S}$ consists precisely of the charts compatible with all charts in $\mathcal{A}$.


\end{exercise}




\begin{exercise}
Let $M$ be a smooth manifold and $\pi : \tilde{M} \to M$ a covering map with $\tilde{M}$ second countable. Show that $\tilde{M}$ admits a smooth structure for which $\pi$ is smooth. 
\begin{hint}
    Consider a chart $(U, \varphi)$ of $M$ with $U \subset M $ an evenly covered neighborhood. For each section $ \sigma :  U \to \tilde{M} $, consider the chart $(\sigma (U) , \varphi \circ \pi  )$. Show that the set of charts of $\tilde{M}$ obtained this way is a smooth atlas. Recall that a section is a continuous map $\sigma : U \to \tilde{M}$ with $\pi \circ \sigma = \text{Id}_U$.
\end{hint}
\end{exercise}



\begin{exercise}
Let $M_n (\mathbb{R})$ denote the space of $n\times n$ real matrices and identify it with $\mathbb{R}^{n^2}$.    Show that 
    \[      SL (n ; \mathbb{R} ) : = \{ A \in M_n (\mathbb{R}) \, \vert \, \det (A) = 1 \}         \]
    is a smooth manifold of dimension $n^2 -1 $. 
    %\begin{hint}
    %    There is a three-line solution. 
    %\end{hint}
\end{exercise}


\begin{exercise}
    Given $A \in M_n(\mathbb{R})$, we denote by $A^T $ its transpose. Show that
    \[      O  (n ) : = \{ A \in M_n (\mathbb{R}) \, \vert \, A A^{T}  = \text{Id} \}         \]
    is a smooth manifold of dimension $n(n-1)/2$. 

    \begin{hint}
        Identify $M_n (\mathbb{R}) = \mathbb{R}^{n^2}$ with $(\mathbb{R}^n)^n$, the set of $n$-tuples of elements of $\mathbb{R}^n$, and $\mathbb{R}^{n(n+1)/2}$ with the set of upper-triangular $n \times n$ matrices.   Define 
        \[    F :    (\mathbb{R}^n)^n \to \mathbb{R}^{n(n+1)/2}      \]
        by
        \[      F _{ij}(  \Vec{v}_1, \ldots,  \Vec{v}_n ) = \Vec{v}_i \cdot \Vec{v}_j - \delta _{ij}                 \]
        for $1\leq i \leq j\leq n$, where $\delta_{ij}$ denotes the Kronecker delta. Show that 
        \[        F^{-1} (0) = O (n)                      \]
        and that for all  $  p \in O (n)    $ the differential $d_p F :  (\mathbb{R}^n)^n \to \mathbb{R}^{n(n+1)/2} $ is surjective. 
    \end{hint}
\end{exercise}










\end{document}